---
title: "Analysis"
author: "Kate Jones, Gianni Spiga"
date: "2022-11-22"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE 
    theme: flatly
    df_print: paged
    
---


# Libraries
```{r, message = FALSE}
library(ggplot2)
library(GGally)
library(plotly)
library(ggcorrplot)
library(car)
```

# Loading in Data
```{r}
aba <- read.table("abalone.txt", sep = ",")
# Fixing Column names 
names(aba) <-
  c(
    "Sex",
    "Length",
    "Diameter",
    "Height",
    "Whole.weight",
    "Shucked.weight",
    "Viscera.weight",
    "Shell.weight",
    "Rings"
  )
head(aba, 15)

# Change Sex into factor 
aba$Sex <- as.factor(aba$Sex)

# Search for NA values
any(is.na(aba))
```

# Visualizations
```{r}
#hist(aba$Length)

# Histrogram of Length
ggplot(data = aba, aes(x = Length)) + geom_histogram(fill = "darkseagreen1", )
mean(aba$Length)
median(aba$Length)

ggplot(data = aba, aes(x = Diameter)) + geom_histogram(fill = "darkseagreen2")
mean(aba$Diameter)
median(aba$Diameter)

ggplot(data = aba, aes(x = Height)) + geom_histogram(fill = "darkseagreen3", bins=100)
mean(aba$Height)
median(aba$Height)

ggplot(data = aba, aes(x = Whole.weight)) + geom_histogram(fill = "darkseagreen4")
mean(aba$Whole.weight)
median(aba$Whole.weight)

ggplot(data = aba, aes(x = Shucked.weight)) + geom_histogram(fill = "darkseagreen")
mean(aba$Shucked.weight)
median(aba$Shucked.weight)

ggplot(data = aba, aes(x = Viscera.weight)) + geom_histogram(fill = "darkseagreen1")
mean(aba$Viscera.weight)
median(aba$Viscera.weight)

ggplot(data = aba, aes(x = Shell.weight)) + geom_histogram(fill = "darkseagreen2")
mean(aba$Shell.weight)
median(aba$Shell.weight)

ggplot(data = aba, aes(x = Rings)) + geom_histogram(bins = 29, fill = "darkseagreen3")
mean(aba$Rings)
median(aba$Rings)

# Scatter plot of Diameter, Rings, and Sex with plotly
ggplotly(ggplot(data = aba, aes(
  x = Diameter, y = Rings, color = Sex
)) + geom_point() + scale_color_manual(values = c("#f8674f", "#2887c8", "#9fe2bf"))) 

# For pie chart 
colors <- c("#f8674f", "#2887c8", "#9fe2bf")

fig1 <-
  plot_ly(
    data = data.frame(table(aba$Sex)),
    labels = ~ Var1,
    values = ~ Freq,
    type = 'pie',
    marker = list(
      colors = colors,
      line = list(color = '#FFFFFF', width = 1.5)
    )
  ) %>% layout(title = "Pie Chart of Abalone Sex")
fig1
```
# Checking relationships amongst variables 

```{r}
corMat <- cor(aba[, -1])
corMat

# heatPlot <- plot_ly(z = corMat, type = "heatmap")
# heatPlot
ggcorrplot(
  corMat,
  type = "lower",
  outline.col = "white",
  lab = TRUE,
  color = c("#f8674f", "#2887c8", "#9fe2bf")
)

# Lots of high multicollinearity!! 
```

```{r}
# Initial Model 
fullMod <- lm(Rings ~ ., data = aba)
summary(fullMod)
anova(fullMod)
vif(fullMod)
```

The VIF indicates that we have high multicollinearity and we will need to build a model accordingly. We now check if our response variable needs a transformation.

```{r}
### Checking Assumptions
plot(fullMod)

# Check for transformation 
library(MASS)
boxcox(fullMod)
vif(fullMod)

# Now we use full log model
logFullMod <- lm(log(Rings) ~ ., data = aba)
summary(logFullMod)
plot(logFullMod)
```

## Addressing Outliers

```{r}
aba[2050:2055,]
#2052, despite its size being similar to other abalones, has an abnormally large weight and height

# aba[160:166, ]
# OBS 164 is an abnormally heavy large female abalone

aba[235:240, ]
# Obs 237 is an extremely young and tiny infant 


# We will drop observations 237 and 2052 since it is harming our assumption of normality in residuals 
aba <- aba[-c(237, 2052),]
```

# Addressing Multicollinearity

```{r}
# Remove Shucked weight since full weight would be easier process
# noShuckMod <- lm(log(Rings) ~ .-Shucked.weight, data = aba)
# summary(noShuckMod)

# Male sex is becoming obviously insignifcant, so lets make new column where abalone is infant or not
aba$Infant <- as.factor(ifelse(aba$Sex == 'I', 'Y','N'))
aba$Infant

# Clean model by removing Length, and Viscera weight
redMod <- lm(log(Rings) ~ Infant + Diameter + Height+ Whole.weight, data = aba)
summary(redMod)
```

```{r}
# Model suggested by ridge regression in mine.RMD, took out whole weight
ridgeSug <- lm(log(Rings) ~ Infant + Length +  Diameter + Height+ Shucked.weight + Viscera.weight + Shell.weight, data = aba)
summary(ridgeSug)

# Lets drop viscera weight 
ridgeSug <- lm(log(Rings) ~ Infant  + Length +  Diameter + Height+ Shucked.weight + Shell.weight, data = aba)
summary(ridgeSug)
```


```{r}
#anova(logFullMod, redMod)

nullModel <- lm(log(Rings) ~ 1, data = aba)
fullModI <- lm(log(Rings) ~ .-Sex, data = aba)


AICstair <-
  stepAIC(
    nullModel,
    scope = list("lower" = nullModel , "upper" = ridgeSug),
    direction = "forward",
    k = 2,
  )

BICstair <-
  stepAIC(
    nullModel,
    scope = list("lower" = nullModel , "upper" = ridgeSug),
    direction = "forward",
    k = log(nrow(aba)),
  )

AICstair
```
```{r}
# We notice that length is the last to be added via step AIC, as well as it is highly correlated with variables (diameter = 0.99), we will drop it 
# BIC also suggests we drop Length 
postStepModel <-
  lm(log(Rings) ~ Infant + Diameter + Height + Shucked.weight + Shell.weight,
     data = aba)
summary(postStepModel)

# Test with stepAIC once more 
AICstair2 <-
  stepAIC(
    nullModel,
    scope = list("lower" = nullModel , "upper" = postStepModel),
    direction = "forward",
    k = 2,
  )

BICstair2 <-
  stepAIC(
    nullModel,
    scope = list("lower" = nullModel , "upper" = postStepModel),
    direction = "forward",
    k = log(nrow(aba)),
  )
```


```{r}
# Checking for outliers
plot(postStepModel)

# aba[2050:2055,]
# #2052, despite its size being similar to other abalones, has an abnormally large weight and height
# 
# aba[160:166, ]
# # OBS 164 is an abnormally heavy large female abalone
# 
# aba[235:240, ]
# # Obs 237 is an extremely young and tiny infant 
```

# Checking for Interactions amongst remianing variables 

```{r}
# Reducing the abalone dataset
aba_red <- aba[,c("Height", "Shucked.weight", "Shell.weight", "Diameter", "Infant", "Rings")]
head(aba_red, 10)

# Now we build a model with all our variabels and interactions
intModel <- lm(log(Rings) ~ .^2, data = aba_red)
summary(intModel)

# Drop all non-signicant variables: height:shucked weight, height:infant, shucked.weight:shellweight, shell.weight:infant

intModel.red <-
  lm(
    log(Rings) ~ Height + Shucked.weight + Shell.weight + Diameter + Infant + Height:Diameter + Shucked.weight:Diameter + Shucked.weight:Infant + Diameter:Infant,
    data = aba_red
  )
summary(intModel.red)

# Removing non significant variable Diameter:Infant 
intModel.red <-
  lm(
    log(Rings) ~ Height + Shucked.weight + Shell.weight + Diameter + Infant + Height:Infant + Height:Diameter + Shucked.weight:Diameter + Shucked.weight:Infant,
    data = aba_red
  )
summary(intModel.red)

cat("\n Checking AIC now \n")
# AIC step with interactions 
AICstair3 <-
  stepAIC(
    nullModel,
    scope = list("lower" = nullModel , "upper" = intModel.red),
    direction = "forward",
    k = 2,
  )

cat("\n Checking BIC now \n")
# BICstepwise with int 
BICstair3 <-
  stepAIC(
    nullModel,
    scope = list("lower" = nullModel , "upper" = intModel.red),
    direction = "forward",
    k = log(nrow(aba)),
  )


# BIC suggests we drop interaction between height:infant 
intModel.final <-
  lm(
    log(Rings) ~ Height + Shucked.weight + Shell.weight + Diameter + Infant+ Height:Diameter + Shucked.weight:Diameter + Shucked.weight:Infant,
    data = aba_red
  )
summary(intModel.final)
```